---
title: "WiDS exploratory data analysis"
author: "Yihan Shi"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, 
                      message = FALSE, warning = FALSE)
```

Diabetes risk factors: 
- pregnant women over 35 years old
- age (in general)
- overweight (BMI >= 25)
- family history
- history of polycystic ovary syndrome (PCOS)
- Injury to the pancreas 
- Autoimmune disease
- Physical stress (such as surgery or illness)
- High blood pressure
- Abnormal blood cholesterol or triglyceride levels
- Smoking


Variables that potentially contribute to higher possibiltiy of diabetes:
1. BMI higher than 25 (overweight)
2. A high ratio (4) between systolic pressure (d1_sysbp_max) over diastolic
pressure (d1_diasbp_max)
3. A high glucose level (d1_glucose_max/d1_glucose_min) - 11.1~33.3 mmol/L 

Target variable: diabetes_mellitus

https://care.diabetesjournals.org/content/31/Supplement_2/S249#:~:text=Baseline%20blood%20pressure%20data%20from,of%20%3C130%2F80%20mmHg.
http://www.diabetesincontrol.com/diabetes-mellitus-and-infectious-diseases-controlling-chronic-hyperglycemia/#:~:text=Patients%20with%20uncontrolled%20diabetes%20are,overall%20immunity%20through%20different%20mechanisms.
https://my.clevelandclinic.org/health/diseases/7104-diabetes-mellitus-an-overview

## Load packages and data

```{r packages}
library(tidyverse)
library(broom)
library(class)
library(infer)
library(dplyr)
library(usethis)
training_full <- read_csv("~/Desktop/Duke 2021-22/WiDs/widsdatathon2021/TrainingWiDS2021.csv")
unlabeled <- read_csv("~/Desktop/Duke 2021-22/WiDs/widsdatathon2021/UnlabeledWiDS2021.csv")
```


```{r cleaning}
training_1 <- training_full%>% 
  select(encounter_id, age, gender, bmi, 
          d1_sysbp_max, d1_sysbp_min,
          d1_diasbp_max, d1_diasbp_min,
          d1_glucose_max, d1_glucose_min,
          cirrhosis, immunosuppression,
          elective_surgery,
          diabetes_mellitus) %>% 
  na.omit(encounter_id, age, gender, bmi, 
          d1_sysbp_max, d1_sysbp_min,
          d1_diasbp_max, d1_diasbp_min,
          d1_glucose_max, d1_glucose_min,
          cirrhosis, immunosuppression,
          elective_surgery,
          diabetes_mellitus)

unlabeled <- unlabeled %>% 
  na.omit(cols = "d1_sysbp_max", "d1_diasbp_max") %>% 
  mutate(SD_pressure_ratio = d1_sysbp_max / d1_diasbp_max)

training_1 <- training_1 %>% 
  group_by(diabetes_mellitus) %>% 
  mutate(SD_pressure_ratio = d1_sysbp_max / d1_diasbp_max,
        glucose_avg = (d1_glucose_max + d1_glucose_min) / 2,
        meanbmi = mean(bmi))
```

Step 1: Simple linear regression test

```{r test BMI}
ggplot(data = training_1, mapping = aes(x = bmi)) + 
  geom_histogram() + 
  labs(title = "Distributio of BMI")

training_1 %>% 
  group_by(diabetes_mellitus) %>% 
  summarise(meanbmi = mean(bmi),
            median = median(bmi),
            IQR = IQR(bmi),
            q1 = quantile(bmi, 0.25),
            q3 = quantile(bmi, 0.75),
            sd = sd(bmi))

ggplot(data = training_1, mapping = aes(x = factor(diabetes_mellitus), y = bmi)) + 
  geom_boxplot() + 
  labs(title = "Mean BMI difference between 2 groups")
```
BMI is right skewed with a median of 27.8, which is higher than 25, the benchmark
of overweight. 50% of the observations fall between 23.7 and 33.

From the boxplot, there is little difference between the 2 groups. However, since
the histogram is right skewed, we use median. For patients who have diabetes, the 
median BMI is 30.3. For patients who do not have diabetes, the median BMI is 27.2. 

```{r BMI linear model}
bmi_model <-  lm(bmi ~ factor(diabetes_mellitus), data = training_1)
tidy(bmi_model)

glance(bmi_model) %>% 
  pull(r.squared)
```
```{r}
mean_diff <- training_1 %>%
  count(diabetes_mellitus, meanbmi) %>%
  select(meanbmi) %>% 
  pull() %>%
  diff()
mean_diff
```


The intercept interprets that a patient who does not have diabetes will have a 
BMI of 28.6. If a patient has diabetes, then the BMI will increases by 3.32.

However, the r squared value is 0.028 only 2.82% of the variability in BMI can be 
explained by diabetes indicator.

Next, we include SD_pressure_ratio and glucose level in the model, respectively.

```{r SD_pressure_ratio}
BP_model <-  lm(SD_pressure_ratio ~ factor(diabetes_mellitus), data = training_1)
tidy(BP_model)

glance(BP_model) %>% 
  pull(r.squared)
```
The intercept interprets that a patient who does not have diabetes will have a 
D/S ratio of 1.7. If a patient has diabetes, then the ratio will increases by 0.08.


```{r glucose level}
glucose_model <-  lm(d1_glucose_max ~ factor(diabetes_mellitus), data = training_1)
tidy(glucose_model)

glance(glucose_model) %>% 
  pull(r.squared)
```
(I doubt that the glucose level unit in the dataset is wrong... Need double check.
In the codebook, the unit is mmol/L, while the actual data shoudl be in mg/dL?)

The intercept interprets that a patient who does not have diabetes will have a 
maximum glucose level of of 156.0. If a patient has diabetes, then the maximum 
glucose level will increases by 83.

16.1% of the variability in maximum glucose level can be explained by diabetes 
indicator.


Step 2: Logistic regression test

```{r create test set}
set.seed(0219)
indices <- sample(nrow(training_1), 250)
```

```{r}
train <- training_1 %>%
  slice(-indices) %>% 
  select(diabetes_mellitus, elective_surgery, SD_pressure_ratio, d1_glucose_max)

test <- training_1 %>%
  slice(indices) %>% 
  select(diabetes_mellitus, elective_surgery, SD_pressure_ratio, d1_glucose_max)

train_status <- training_1 %>%
  slice(-indices) %>% 
  pull(diabetes_mellitus)

unlabeled_result <- unlabeled %>% 
  select(elective_surgery, SD_pressure_ratio, d1_glucose_max)
```

```{r}
true_status <- training_1 %>%
  slice(indices) %>% 
  pull(diabetes_mellitus)
```

```{r knn}
mod_knn <- knn(train, unlabeled, train_status, 
               k = 10, prob = F, use.all = F)
mod_knn == true_status
mean(mod_knn == true_status)
```
99.7% of the test data was accurately classified.

```{r logistic model}
train <- training_1 %>%
  slice(-indices) %>% 
  select(diabetes_mellitus, bmi, SD_pressure_ratio, d1_glucose_max)

train <- training_1 %>% 
  mutate(diabetes_bin = dplyr::if_else(diabetes_mellitus == "Yes", 1, 0))

logit_mod <- glm(diabetes_bin ~ factor(elective_surgery) + SD_pressure_ratio + d1_glucose_max,
                 data = train, family = "binomial")

tidy(logit_mod) %>% 
  select(term, estimate)
```


```{r prediction accuracy}
pred_log_odds <- augment(logit_mod, newdata = test) %>% 
  pull(.fitted)

pred_probs <- exp(pred_log_odds) / (1 + exp(pred_log_odds))
round(pred_probs, 3)

true_status

breaks <- c(0, .5, 1)
tags <- c(0,1)
pred_status <- cut(pred_probs, 
                  breaks=breaks, 
                  include.lowest=TRUE, 
                  right=FALSE, 
                  labels=tags)
pred_status
mean(pred_status == true_status)
```
The accuracy of the logistic model is 77.6%.

Using BMI, SD ratio, and maximum glucose level, the k-NN model has an accuracy of
95.6%, while the logistic model has an accuracy of 77.6%. 

 